---
title: "covid 19"
author: "Alejandra Magdaleno"
date: "November 9, 2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```
  
This graph shows that there was a larger number of females than males in the alive category.  It also shows that there was a larger number of males than females in the dead category. 
  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata,type = "percent", main ="Frequency Bar Chart")
```

This graph displays the information in percentages.  It shows that out of those that are alive, approximately 59% are females and 38% are males.  Of those that are dead, approximately 38% are females and 62% are males.  If dying is independent of sex, it is unexpected to see a much larger percentage of males dead as opposed to females.       


  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

The table shows that out the people who were alive, 56% were female and 44% were males.  Of those that were dead, 36% were female and 64% were males.  This is consistent with our previous observation that males have a higher frequency of dying that females do.  

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Fisher's Exact Test results reflects a low p-value and a 95% confidence level between 1.4 and 3.9.  The odds ratio of 2.3 to 1 shows that males are 2.3 times more likely to die than females as a result of covid. 

## Conclusions

 The graphical, numerical, and inferential results show that males are more likely to die than females as a result of covid19.  The results of the Fisher's Exact Test shows that Males are 2.3 times more likely to die than females.  As a result, we will reject the null hypothesis stating that men and women are equally likely to die of covid19.   